Evaluating a simplified forecast model in comparison to the ECDC forecasting hub ensemble

### 2022-08-24 paper/paper.Rmd
update publishing

The model is implemented in `stan` [@stan] and `R` (`4.2.0`) [@R] as an extension of the baseline model from the `forecast.vocs` R package (`0.0.9.7000`) [@forecast.vocs].
The `cmdstanr` R package (`0.5.2`) [@cmdstanr] is used for model fitting with 2 MCMC chains each having 1000 warm-up and 1000 sampling steps each [@cmdstanr].
`cmstanr` surfaces several settings which trade-off between sampling speed and the robustness of approach.
Here we take a conservative approach, as model fit is not manually inspected during real-time usage and due to the expected complexity of the posterior [@betancourt_2017], and set the adapt delta setting to 0.99, and the maximum treedepth setting to 15. 
For real-time usage convergance was not assessed but during model development the Rhat diagnostic was used alongside feedback from `cmdstanr` about the number of divergant transitions and exceedance of the maximum tree depth. During development posterior predictions were also visually compared to observed data [@cmdstanr].

@@

In this study, we defined a surrogate model aiming to replicate some of the observed behaviour of the ECDC forecast hub multi-team ensemble for forecasting test positive reported COVID-19 cases in ECDC nations and the UK. We first defined a set of assumptions for how the surrogate model should behaved based on our observations of the ECDC forecast hub ensemble, our experiance submitting forecasts to various forecast Hubs and our observations of the performance of other models. We restricted ourselves to a model that could be easily understood, that produced meaningful summary statistics, and that could be run with low compute resources. We further provide a fully reproducibile workflow for running this model using GitHub actions and code to evaluate its performance.

>>>>

The model is implemented in `stan` [@stan] and `R` (`4.2.0`) [@R] as an extension of the baseline model from the `forecast.vocs` R package (`0.0.9.7000`) [@forecast.vocs].
We note that our use of a Heaviside step function introduces a discontinuity to the posterior making it less suited for use with `stan`. Other model formulations without this would be more efficient.
The `cmdstanr` R package (`0.5.2`) [@cmdstanr] is used for model fitting with 2 MCMC chains each having 1000 warm-up and 1000 sampling steps each [@cmdstanr].
`cmstanr` surfaces several settings which trade-off between sampling speed and the robustness of approach.
Here we take a conservative approach, as model fit is not manually inspected during real-time usage and due to the expected complexity of the posterior [@betancourt_2017], and set the adapt delta setting to 0.99, and the maximum treedepth setting to 15. 
For real-time usage convergance was not assessed but during model development the Rhat diagnostic was used alongside feedback from `cmdstanr` about the number of divergant transitions and exceedance of the maximum tree depth. During development posterior predictions were also visually compared to observed data [@cmdstanr].

@@

In this study, we defined a surrogate model aiming to replicate some of the observed behaviour of the ECDC forecast hub multi-team ensemble for forecasting test positive reported COVID-19 cases in ECDC nations and the UK. We first defined a set of assumptions for how the surrogate model should behaved based on our observations of the ECDC forecast hub ensemble, our experiance submitting forecasts to various forecast Hubs, and our observations of the performance of other models. We restricted ourselves to a model that could be easily understood, that produced epidemiologically meaningful summary statistics, and that could be run with low compute resources. We further provide a fully reproducibile workflow for running this model using GitHub actions and code to evaluate its performance.
